An Executable Formal Semantics For A Functional Actor Language

### 2017-12-20 paper/paper.tex
nothing
add sec 2.3 & 2.4

\section{The Actor Language in K}
In this section, we describe in detail some important features of the actor
language from~\cite{}. We cover some important features from K and how they were
used to define key concepts of the actor language.

@@

\subsection{Semantics}\label{sec:semantics}

\subsection{In-depth Example}

>>>>

\section{The Actor Language in K}
In this section, we describe in detail some important features of the actor
language from~\cite{actor}. We cover some important features from K and how 
they were
used to define key concepts of the actor language.

@@

\subsection{Semantics}\label{sec:semantics}
Semantics in K is given as \emph{rewriting rules} of the form $l \Rightarrow r$.
For example, the  beta-reduction rule (beta-v) is specified as
\begin{verbatim}
  rule app(lambda(X:Id, E:Exp), V:Val) => E [ V / X ]
\end{verbatim}
where \texttt{E [ V / X ]} is the result of substituting \texttt{V} for 
\texttt{X} in \texttt{E} using K's internal substitution mechanism.
Conditional reduction rules can also be easily specified.
For example, the branching rule (br) is specified as two K rules
\begin{verbatim}
  rule br(nil, V1:Val, V2:Val) => V2
  rule <k> br(V:Val, V1:Val, V2:Val) => V1 ... </k>
       <definedaddr> X </definedaddr>
    requires notBool((isId(V) andBool notBool(V in X)) orBool V ==K nil)
\end{verbatim}
The first rule says if the first argument is \texttt{nil}, the branching 
operator \texttt{br} returns its third argument.
The second rule is a bit complicated.
It says if the first argument is neither \texttt{nil} nor 
unknown addresses, the branching operator \texttt{br} returns its second 
argument.

The single-step transition relation $\mapsto$ on actor configurations is 
similarly specified using K rules.
For example, the \texttt{<init: a, a1>} transition rule is specified as
\begin{verbatim}
  rule <actor>
         <k> initbeh(A1, V:Val) => nil ... </k>
         <id> A </id>
       </actor>
       <actor>
         <k> uninit(A) => ready(V) ... </k>
         <id> A1 </id>
       </actor>
\end{verbatim}
The rule involves two actors as there are two actor cells 
\texttt{<actor>}\ \dots\ \texttt{</actor>} in the rule.
The first actor is the \emph{initializer} actor while the second is the actor 
\emph{being initialized}.
K supports to only specify components that are relevant to the rule. This is 
known as the \emph{framing principle}.
The initializer actor executes \texttt{initbeh(A1, V:Val)} and gets value 
\texttt{nil}, as specified in the rule.
The rest computation or \emph{the reduction context} is unchanged and kept in 
the following~``\texttt{...}''.

Reduction context can be mentioned explicitly when needed.
For example in the following \texttt{<bec: a, a1>} rule,
\begin{verbatim}
  rule <actor>
         <k> (become(V:Val) ~> RestComputation) => ready(V) </k>
         ...
       </actor>
       (.Bag => <actor>
                  <k> nil ~> RestComputation </k>
                  <id> !A1:Id </id>
                </actor>)
       <definedaddr> Ads => Ads ListItem(!A1) </definedaddr>
\end{verbatim}
the reduction context \texttt{RestComputation} is explicitly passed to a 
newly-generated actor.
The tilde arrow expression \verb|T ~> R| 
means the decomposition \verb|R[T]| with reduction context \texttt{R} and redex 
\texttt{T}.

\subsection{In-depth Example}
We test our actor semantics on an example of tree recursion~\cite{actor}.
The problem is to calculate the product of all leaves of a binary tree.
The problem is solved by a divide-and-conquer algorithm which calculates the 
results for two subtrees and multiplies the results.
The algorithm is implemented using a \emph{join continuation} as follows.
\begin{verbatim}
let(bind(sink,
  rec(lambda(b, lambda(m, become(b))))),
let(bind(ident,
  lambda(m, m)),
let(bind(joincont,
  rec(lambda(b, lambda(c, lambda(nargs, lambda(firstnum, lambda(num,
    if(eq?(nargs, 0),
       become(app(app(app(b, c), 1), num)),
       seq(become(sink),
           send(c, mult(firstnum, num))))))))))),
let(bind(treeprod,
  rec(lambda(b, lambda(self, lambda(m,
    seq(become(app(b, self)),
        if(isnat(tree(m)),
           send(cust(m), tree(m)),
           letactor(bind(newcust, app(app(app(joincont, cust(m)), 0), nil)),
             seq(send(self, pr(left(tree(m)), newcust)),
                 send(self, pr(right(tree(m)), newcust))))))))))),
let(bind(tr, 
  pr(pr(1, pr(2, pr(3, 4))), pr(pr(5, 6), 7))),
letactor(
  bind(tpactor, app(treeprod, tpactor)),
  bind(custactor, ident),
  send(tpactor, pr(tr, custactor))))))))
\end{verbatim}
Remind that our semantics initializes the initial actor with the behavior of 
executing the above lambda expression.
Therefore after five \texttt{let}-expressions, the initial actor creates two 
actors. One is the \texttt{tpactor} that calculates the tree product.
The other is the \texttt{custactor} that receives the result.
The initial actor sends \texttt{tpactor} a message containing a binary tree 
whose leaves are $1, 2, \dots, 7$, and the address of the customer actor 
\texttt{custactor}.
When the actor system terminates, we get the following configuration
\begin{verbatim}
<T>
  <actors>
    <actor>
      <k> ready ( ... ) </k>
      <id> _93 </id>
    </actor>
    <actor>
      <k> ready ( ... ) </k>
      <id> _179 </id>
    </actor>
    .... ....
    <actor>
      <k> ready ( ... ) </k>
      <id> _45 </id>
    </actor>
    <actor>
      <k> exec ( nil ) </k>
      <id> initactor </id>
    </actor>
    <actor>
      <k> exec ( 5040 ) </k>
      <id> _46 </id>
    </actor>
    ......
    ......
  </actors>
  <messages> .MessageCellBag </messages>
  <definedaddr> ... </definedaddr>
</T>
\end{verbatim}
In the configuration there are many idle actors generated during the 
computation. The customer actor, which in this execution has identity 
\texttt{\_46}, is holding the final product of 5040.
